---
title: "tibbles.practice"
author: "Veronica"
date: "5/29/2017"
output: 
  html_document: 
    keep_md: yes
    self_contained: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nycflights13)
```

### 10. Tibbles

#### Exercises  

**1. How can you tell if an object is a tibble? (Hint: try printing mtcars, which is a regular data frame).**  
```{r}
is.tibble(mtcars)
print(mtcars)

is.tibble(diamonds)
print(diamonds)
```  

is.tibble() will print TRUE if a data set is a tibble and FALSE if not. A tibble will only print the first 10 rows in the console, while a data frame will print the entire data set.

**2. Compare and contrast the following operations on a data.frame and equivalent tibble. What is different? Why might the default data frame behaviours cause you frustration?**  
```{r}
df <- data.frame(abc = 1, xyz = "a")
tib <- as_tibble(df)

df$x
tib$x
#the tibble does not return a value because the match is not identical to the variable name.

df[, "xyz"]
tib[, "xyz"]
#the tibble returns the values in a tibble

df[, c("abc", "xyz")]
tib[, c("abc", "xyz")]
#the tible includes the variable type.
```

**3. If you have the name of a variable stored in an object, e.g. var <- "mpg", how can you extract the reference variable from a tibble?**  
```{r}
var <- "mpg"

mtcars %>% 
  as_tibble() %>% 
  .[[var]]

mtcars %>% 
  as_tibble() %>% 
  .$mpg

mtcars %>% 
  as_tibble() %>% 
  .[["mpg"]]
```  
$ extracts by name and is not compatable with stored variables. .[[var]] will extract the variable stored in var, but if searching for a variable by name quotes must be used.

**4. Practice referring to non-syntactic names in the following data frame by:**  
```{r}
annoying <- tibble(
  `1` = 1:10,
  `2` = `1` * 2 + rnorm(length(`1`))
)

#Extracting the variable called 1.
annoying[[1]]

#Plotting a scatterplot of 1 vs 2.
ggplot(annoying, aes(`1`, `2`)) +
  geom_point()

#Creating a new column called 3 which is 2 divided by 1.
annoying2 <- mutate(annoying, `3` = `2`/`1`)

#Renaming the columns to one, two and three.
annoying2 %>% 
  rename(
    one = `1`,
    two = `2`,
    three = `3`
  )
```  

**5. What does tibble::enframe() do? When might you use it?**  
```{r}
x <- (letters) %>% 
  enframe(name = "red", value = "blue") %>% 
  print()
```  
enframe() coerces a vector into a tibble. The position is listed under "name" and the value under "value". Arguments to rename the columns can be passed in enframe().

**6. What option controls how many additional column names are printed at the footer of a tibble?**  
```{r}
print(flights)
options(tibble.max_extra_cols = 3)
print (flights)
```  





