---
title: "Data.Transformation.1"
author: "Veronica"
date: "5/7/2017"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nycflights13)
```
### Data Transformation

#### 5.2 Filter rows with filter()

**1. Find all flights that for the following**  
Had an arrival delay of two or more hours  
```{r}
filter(flights, arr_delay >= 120)
```

Flew to Houston (IAH or HOU)  
```{r}
filter(flights, dest == "IAH" | dest == "HOU")
```

Were operated by United, American, or Delta
```{r}
filter(flights, carrier %in% c("UA", "AA", "DL"))
```

Departed in summer (July, August, and September)
```{r}
filter(flights, month %in% c(7:9))
```

Arrived more than two hours late, but didn’t leave late
```{r}
filter(flights, dep_delay <=0 & arr_delay >= 120)
```

Were delayed by at least an hour, but made up over 30 minutes in flight
```{r}
filter(flights, dep_delay >= 60 & dep_delay-arr_delay >=30)
```

Departed between midnight and 6am (inclusive)
```{r}
filter(flights, dep_time %in% c(0000:0600))
```

**2. Another useful dplyr filtering helper is between(). What does it do? Can you use it to simplify the code needed to answer the previous challenges?**  
between() filters data by variables that fall between two values. if could have been used to filter flights that left between midnight and 6am, or summer flights.

**3. How many flights have a missing dep_time? What other variables are missing? What might these rows represent?** 
```{r}
filter(flights, is.na(dep_time))
```  
8,255 flights have NA in the depature time variable. Departure delay, arrival time ,arrival delay and  air time can have missing variables. The flights with these missing values were likely canceled. 

**4. Why is NA ^ 0 not missing? Why is NA | TRUE not missing? Why is FALSE & NA not missing? Can you figure out the general rule? (NA * 0 is a tricky counterexample!)**  


#### 5.3 Arrange rows with arrange()

**1. How could you use arrange() to sort all missing values to the start? (Hint: use is.na())**  
```{r}
arrange(flights, desc(is.na(dep_time), is.na(arr_time)))
```

**2. Sort flights to find the most delayed flights. Find the flights that left earliest.**  
```{r}
arrange(flights , desc(dep_delay))
arrange(flights, dep_delay)
```

**3. Sort flights to find the fastest flights.**
```{r}
arrange(flights, air_time)
```

**4. Which flights travelled the longest? Which travelled the shortest?**
```{r}
arrange(flights, desc(distance))
arrange(flights, distance)
```

#### 5.4 Select columns with select()

**1. Brainstorm as many ways as possible to select dep_time, dep_delay, arr_time, and arr_delay from flights.**
```{r}
select(flights, dep_time, dep_delay, arr_time, arr_delay)
select(flights, 4, 6, 7, 9)
select(flights, dep_time, arr_time, ends_with("delay"))
```

**2. What happens if you include the name of a variable multiple times in a select() call?**
```{r}
select(flights, arr_time, arr_time)
```  
The resulting data frame (or tibble) will have the called variable only once.

**3. What does the one_of() function do? Why might it be helpful in conjunction with this vector?**
```{r}
vars <- c("year", "month", "day", "dep_delay", "arr_delay")
select(flights, one_of(vars))
```
one_of() is a select helper that allows you to select all of the variables included in a character vector. 

**4. Does the result of running the following code surprise you? How do the select helpers deal with case by default? How can you change that default?**  
```{r}
select(flights, contains("TIME"))
```
select() ignores case of the arguments by default. To chnage the default ignore.case = FALSE must be passed as an argument

#### 5.5 Add new variables with mutate()

**1. Currently dep_time and sched_dep_time are convenient to look at, but hard to compute with because they’re not really continuous numbers. Convert them to a more convenient representation of number of minutes since midnight.**  
hour = dep_time %/% 100,
  minute = dep_time %% 100
```{r}
flights2 <- mutate(flights, 
       dep_time_min = dep_time %/% 100 *60 + dep_time %% 100,
       sched_dep_time_min = sched_dep_time %/% 100 *60 + sched_dep_time %% 100)
select(flights2, dep_time, dep_time_min, sched_dep_time, sched_dep_time_min)
```

**2. Compare air_time with arr_time - dep_time. What do you expect to see? What do you see? What do you need to do to fix it?**  
```{r}
flights3 <- mutate(flights, arr_diff = arr_time -dep_time)
select(flights3, air_time, arr_diff)
```  
The two values should be the same, however arr_time and dep_time are not listed in minutes past midnight and cannot be used to calculate air_time in this way. Instead, use the departure and arrival times in minutes past midnight as calculated in the first problem.  
```{r}
flights4 <- mutate(flights, 
       dep_time_min = dep_time %/% 100 *60 + dep_time %% 100,
       arr_time_min = arr_time %/% 100 *60 + arr_time %% 100, 
       diff = arr_time_min - dep_time_min)
select(flights4, air_time, diff)       
```  
Not sure why thse still don't match....

**3. Compare dep_time, sched_dep_time, and dep_delay. How would you expect those three numbers to be related?**  
```{r}
select(flights, dep_time : dep_delay)
```
Dep_delay should be the difference between departure time and scheduled departure time. This is trues for differences of less than one hour. dep_time and scheduled departure time are formatted such that subtracting the two will not give the correct value for differences over an hour. Instead, minutes past midnight format should be used to do this calculation.

**4. Find the 10 most delayed flights using a ranking function. How do you want to handle ties? Carefully read the documentation for min_rank().**
```{r}
(x <- arrange(flights, min_rank(desc(dep_delay))))
```
the ties method is "min" in rank_min by default. 

**5. What does 1:3 + 1:10 return? Why?**  
```{r}
1:3 + 1:10
```
The error is returned because the vectors are not the same length, and the longer vector length is not divisible by the shorter vector length. Generally the small vctore will repeat until the longer vector is exhausted but an error is returned if the lengths will not be exhautsed at the same time. 

**6. What trigonometric functions does R provide?**  
The trig functions can be found under ?cos().
 



