---
title: "Data.Transformation.1"
author: "Veronica"
date: "5/7/2017"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nycflights13)
```
### Data Transformation

#### 5.2 Filter

**1. Find all flights that for the following**  
Had an arrival delay of two or more hours  
```{r}
filter(flights, arr_delay >= 120)
```

Flew to Houston (IAH or HOU)  
```{r}
filter(flights, dest == "IAH" | dest == "HOU")
```

Were operated by United, American, or Delta
```{r}
filter(flights, carrier %in% c("UA", "AA", "DL"))
```

Departed in summer (July, August, and September)
```{r}
filter(flights, month %in% c(7:9))
```

Arrived more than two hours late, but didn’t leave late
```{r}
filter(flights, dep_delay <=0 & arr_delay >= 120)
```

Were delayed by at least an hour, but made up over 30 minutes in flight
```{r}
filter(flights, dep_delay >= 60 & dep_delay-arr_delay >=30)
```

Departed between midnight and 6am (inclusive)
```{r}
filter(flights, dep_time %in% c(0000:0600))
```

**2. Another useful dplyr filtering helper is between(). What does it do? Can you use it to simplify the code needed to answer the previous challenges?**