---
title: "Data.import"
author: "Veronica"
date: "6/5/2017"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Data Import

#### 11.2 Getting Started Exercises  

**1. What function would you use to read a file where fields were separated with
“|”?**  
read_delim() can be used to read files if delim = "|"
```{r}
read_delim("delim.txt", delim = "|")
```

**2. Apart from file, skip, and comment, what other arguments do read_csv() and read_tsv() have in common?**  
Both commands share all of the same arguments.

**3. What are the most important arguments to read_fwf()?**  
path and col_positions are the most important arguments of read_fwf().

**4. Sometimes strings in a CSV file contain commas. To prevent them from causing problems they need to be surrounded by a quoting character, like " or '. By convention, read_csv() assumes that the quoting character will be ", and if you want to change it you’ll need to use read_delim() instead. What arguments do you need to specify to read the following text into a data frame?**  
```{r}
read_delim(delim = ",", quote = "'" , "x,y\n1,'a,b'")
```

**5. Identify what is wrong with each of the following inline CSV files. What happens when you run the code?**  
```{r}
#read_csv("a,b\n1,2,3\n4,5,6")
# add an c column
read_csv("a,b,c\n1,2,3\n4,5,6")

#read_csv("a,b,c\n1,2\n1,2,3,4")
#whats happening here?
read_csv("a,b,c\n1,2\n1,2,3,4")

#read_csv("a,b\n\"1")
#wrong coltype
read_csv(col_types = "cc", "a,b\n\"1")

#read_csv("a,b\n1,2\na,b")
#not sure what should be changed
read_csv("a,b\n1,2\na,b")

#read_csv("a;b\n1;3")
#wrong delim, use read_csv2()
read_csv2("a;b\n1;3")
```


#### Parsing a Vector Exercises

**1. What are the most important arguments to locale()?**  
The most important arguments for locale() seems dot be date_format, which sets the default input so that it is parsed as a date. Date_names, demical_mark and grouping mark could also be very important if collaborating outside of the US. 

**2. What happens if you try and set decimal_mark and grouping_mark to the same character? What happens to the default value of grouping_mark when you set decimal_mark to “,”? What happens to the default value of decimal_mark when you set the grouping_mark to “.”?**  
```{r}
#locale(demical_mark = ".", groping grouping_mark = ".")
#Setting decimal_mark and grouping_mark to the same character returns an error

locale(decimal_mark = ",")
#The grouping_mark has changed to ".".
default_locale()

locale(grouping_mark = ".")
#The decimal_mark has changed to ",".
default_locale()
```

**3. I didn’t discuss the date_format and time_format options to locale(). What do they do? Construct an example that shows when they might be useful.**  
date_format and time_format change the default formats for their respective categories.  
```{r}
x <- c("02/03/04", "12/11/10")

#parse_date(x)
#parsing fails

parse_date(x, locale = locale(date_format = "%m/%d/%y"))
parse_date(x, locale = locale(date_format = "%y/%m/%d"))
```

**4. If you live outside the US, create a new locale object that encapsulates the settings for the types of file you read most commonly.**
```{r}
locale("de", decimal_mark = ",", grouping_mark = " ")
n <- c("100 000,3", "300 001,3", "444 444 444,4")
parse_number(n, locale = locale("de", decimal_mark = ",", grouping_mark = " "))
```

**5. What’s the difference between read_csv() and read_csv2()?**  
read_csv() reads files here the feild separator is "," while read_csv2() read files that use ";"

**6. What are the most common encodings used in Europe? What are the most common encodings used in Asia? Do some googling to find out.**  
ISO 8859 seems to have individual encodings for most of the languages in Europe.  
Various chinese sites use GBK, UTF-8, and GB2312. Other Asian websites appear to favor UTF-8.
http://xahlee.info/w/what_encoding_do_chinese_websites_use.html

**7. Generate the correct format string to parse each of the following dates and times:**  
```{r}
d1 <- "January  1, 2010"
d2 <- "2015-Mar-07"
d3 <- "06-Jun-2017"
d4 <- c("August 19 (2015)", "July 1 (2015)")
d5 <- "12/30/14" # Dec 30, 2014
t1 <- "1705"
t2 <- "11:15:10.12 PM"

d1
parse_date(d1, "%B %d, %Y")
d2
parse_date(d2, "%Y-%b-%d")
d3
parse_date(d3, "%d-%b-%Y")
d4
parse_date(d4, "%B %d (%Y)")
d5
parse_date(d5, "%D")
t1
parse_time(t1, "%H%M")
t2
parse_time(t2, "%H:%M:%OS %p")
```












