---
title: "data.transformation.2"
author: "Veronica"
date: "5/15/2017"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nycflights13)
library(lubridate)
```

#### Grouped summaries with summarize()

**1. Brainstorm at least 5 different ways to assess the typical delay characteristics of a group of flights. Consider the following scenarios:**  
*A flight is 15 minutes early 50% of the time, and 15 minutes late 50% of the time.   
*A flight is always 10 minutes late.   
*A flight is 30 minutes early 50% of the time, and 30 minutes late 50% of the time.  
*99% of the time a flight is on time. 1% of the time it’s 2 hours late.  

```{r}
not_cancelled %>%
  group_by(flight) %>%
  summarize(early_15_min = mean(arr_delay == -15), late_15_min = mean(arr_delay == 15)) %>%
  filter(early_15_min == .5 & late_15_min == .5)
```

**2. Come up with another approach that will give you the same output as not_cancelled %>% count(dest) and not_cancelled %>% count(tailnum, wt = distance) (without using count()).**

```{r}
not_cancelled <- flights %>% 
  filter(!is.na(dep_delay), !is.na(arr_delay))
not_cancelled %>% count(dest)

not_cancelled %>% count(tailnum, wt = distance)

not_cancelled %>%
  group_by(dest) %>%
  summarise(flights = n()) 

not_cancelled %>%
  group_by(tailnum) %>%
  summarise(flights = sum(distance))

```

**3. Our definition of cancelled flights (is.na(dep_delay) | is.na(arr_delay) ) is slightly suboptimal. Why? Which is the most important column?**

Flghts that departed but did not arrive indicated that the flights did safely reach their destination. Coding it this way also allows more room for typing errors. dep_delay is more important here, so is.na(dep_delay) can be used instead.

**4. Look at the number of cancelled flights per day. Is there a pattern? Is the proportion of cancelled flights related to the average delay?**   
```{r}
flights %>% 
  mutate(dep_date = lubridate::make_datetime(year, month, day)) %>%
  group_by(dep_date) %>%
  summarize(canceled = sum(is.na(dep_time))) %>%
  ggplot(aes(x = dep_date, y = canceled)) + 
  geom_point() 

flights %>%
  mutate(dep_date = lubridate::make_datetime(year, month, day)) %>%
  group_by(dep_date) %>%
  summarise(proportion_canceled = sum(is.na(dep_time)/n()),
            ave_arr_delay = mean(arr_delay, na.rm = TRUE)) %>%
  ggplot(aes(x = proportion_canceled, y = ave_arr_delay)) + 
  geom_point() +
  geom_smooth()
```  
The proportion of delayed flights and average arival delay are positively correlated.

**5. Which carrier has the worst delays? Challenge: can you disentangle the effects of bad airports vs. bad carriers? Why/why not? (Hint: think about flights %>% group_by(carrier, dest) %>% summarise(n()))**  
```{r}
flights %>% 
  group_by(carrier) %>%
  summarize(ave_arrival_delay = mean(arr_delay, na.rm = TRUE), n = n(), proportion_delay = mean(arr_delay >0, na.rm =TRUE)) %>%
  arrange(desc(proportion_delay))
```  
Frontier Airlines has the longest average arrival delay but a relatively small number of flights compared to most of the carriers. AirTran has the highest proportion of delayed flights.

**What does the sort argument to count() do. When might you use it?**  
The sort argument in count() will arrange the data in descending order based on count(). 

#### 5.7 Group mutates (and filters)

**1. Refer back to the lists of useful mutate and filtering functions. Describe how each operation changes when you combine it with grouping.**  

**2. Which plane (tailnum) has the worst on-time record?**  
```{r}
not_cancelled %>%
  group_by(tailnum) %>%
  summarize(proportion_delay = sum(arr_delay > 0)/n()) %>%
  arrange(desc(proportion_delay))
not_cancelled %>%
  group_by(tailnum) %>%
  summarize(proportion_delay = mean(arr_delay > 0), ave_delay = mean(arr_delay)) %>%
  arrange(desc(proportion_delay))

```
There are many planes that are always delayed.

**3. What time of day should you fly if you want to avoid delays as much as possible?**  
```{r}
not_cancelled %>%
  group_by(sched_dep_time) %>%
  summarize(ave_arr_delay = mean(arr_delay)) %>%
  ggplot(aes(x = sched_dep_time, y = ave_arr_delay)) + 
  geom_point() + 
  geom_smooth()
```  

**4. For each destination, compute the total minutes of delay. For each, flight, compute the proportion of the total delay for its destination.**  
```{r}
not_cancelled %>%
  group_by(dest) %>%
  mutate(dest_delay = sum(arr_delay)) %>%
  group_by(carrier, flight) %>%
  summarize()
  summarize(prop_of_delay = arr_delay/dest_delay)

not_cancelled %>%
  group_by(tailnum, dest) %>%
  summarize(flight_delay = sum(arr_delay))
```
very lost on this one...

**5. Delays are typically temporally correlated: even once the problem that caused the initial delay has been resolved, later flights are delayed to allow earlier flights to leave. Using lag() explore how the delay of a flight is related to the delay of the immediately preceding flight.**  
```{r}
flights %>%
  group_by(year, month, day) %>%
  filter(dest=="LGA", rank(sched_dep_time))
  
```
this is making me cry...

**6. Look at each destination. Can you find flights that are suspiciously fast? (i.e. flights that represent a potential data entry error). Compute the air time a flight relative to the shortest flight to that destination. Which flights were most delayed in the air?**  
```{r}
#flights %>%
#  group_by(dest) %>%
#  summarize(arr_time = )
```  
lost here too...

**7. Find all destinations that are flown by at least two carriers. Use that information to rank the carriers.**
```{r}
flights %>%
  group_by(dest) %>%
  summarize(num_carriers = n_distinct(carrier)) %>%
  filter(num_carriers >= 2) 
```  
Not sure how to rank.

**8. For each plane, count the number of flights before the first delay of greater than 1 hour.**




