---
title: "Tidy.Data1"
author: "Veronica"
date: "6/19/2017"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Tidy Data

#### 12.2 Tidy Data

**1. Using prose, describe how the variables and observations are organised in each of the sample tables.**  
```{r}
table1
#each row contains one obersvation and each column contains one variable.
table2
#this resembles a melted data set. Each row contains one variable from one observation. The columns include "type" or variable, and "count"", or value.
table3
#Each row contains one variable but the variables are combined in one comun such that each cell contains two observed values.
table4a
table4b
#one variable is represented in two columns of each table. There are two observations per row.
```

**2. Compute the rate for table2, and table4a + table4b. You will need to perform four operations:**  
```{r}
t2.1 <- table2 %>% 
  filter(type == "cases")
t2.2 <- table2 %>%
  filter(type == "population")
mutate(t2.1, rate = count / t2.2$count * 10000)

#not sure how to approach the table 4 problem.
```

**3. Recreate the plot showing change in cases over time using table2 instead of table1. What do you need to do first?**
```{r}
table2 %>% 
  filter(type == "cases") %>% 
  ggplot(aes(year, count)) +
  geom_line(aes(group = country), colour = "grey50") + 
  geom_point(aes(colour = country))
#first you must fliter out the cases variable
```

### 12.3 Spreading and gathering

**1. Why are gather() and spread() not perfectly symmetrical? Carefully consider the following example:**  
```{r}
stocks <- tibble(
  year   = c(2015, 2015, 2016, 2016),
  half  = c(   1,    2,     1,    2),
  return = c(1.88, 0.59, 0.92, 0.17)
)
stocks %>% 
  spread(year, return) %>% 
  gather("year", "return", `2015`:`2016`)
```

Gather will only return column names as characters, even if they are numeric 
passing the argument convert  = TRUE requires spread to "guess" what the appropriate value type is.

**2. Why does this code fail?**
```{r}
#table4a %>% 
#  gather(1999, 2000, key = "year", value = "cases")
table4a %>% 
  gather(`1999`, `2000`, key = "year", value = "cases")
```  
The numeric column names require backticks. Here, the columns selected are the 1999 and 2000 coulumns in the data frame.

**3. Why does spreading this tibble fail? How could you add a new column to fix the problem?**  
```{r}
people <- tribble(
  ~name,             ~key,    ~value,
  #-----------------|--------|------
  "Phillip Woods",   "age",       45,
  "Phillip Woods",   "height",   186,
  "Phillip Woods",   "age",       50,
  "Jessica Cordero", "age",       37,
  "Jessica Cordero", "height",   156
)

#people %>% 
#  spread(key = key, value = value)
#Error: Duplicate identifiers for rows (1, 3)

#There is a replication of values for one variable in the same observation
people %>% 
  mutate(ID = c(1:5)) %>% 
  spread(key = key, value = value)
```

**Tidy the simple tibble below. Do you need to spread or gather it? What are the variables?**  
```{r}
preg <- tribble(
  ~pregnant, ~male, ~female,
  "yes",     NA,    10,
  "no",      20,    12
)

preg %>% 
  gather(male, female, key = "gender", value = "count")
```

### 12.4 Separating and Uniting

**1. What do the extra and fill arguments do in separate()? Experiment with the various options for the following two toy datasets.**  
```{r}
tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
  separate(x, c("one", "two", "three"))

tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
  separate(x, c("one", "two", "three"), extra = "drop")

tibble(x = c("a,b,c", "d,e", "f,g,i")) %>% 
  separate(x, c("one", "two", "three"))

tibble(x = c("a,b,c", "d,e", "f,g,i")) %>% 
  separate(x, c("one", "two", "three"), fill = "left", remove = FALSE)
```

**2. Both unite() and separate() have a remove argument. What does it do? Why would you set it to FALSE?**  
```{r}
tibble(x = c("a,b,c", "d,e", "f,g,i")) %>% 
  separate(x, c("one", "two", "three"), fill = "left", remove = FALSE)
```  
The remove arguement, when TRUE, eliminates the original separated column. If there is some abiguity between withing that column, such as missing values, it might be useful to keep the original column by setting remove to FALSE.

**3. Compare and contrast separate() and extract(). Why are there three variations of separation (by position, by separator, and with groups), but only one unite?**  
Extract will isolate an entire column while separate removes two sets of values from a column. 




