---
title: "mpg.practice"
author: "Veronica"
date: "4/24/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## MPG dataset practice
### 3.2 First Steps

```{r}
mpg
```


```{r}
ggplot(data = mpg) +
  geom_point(mapping =aes(x = displ, y = hwy))
```

### 3.2 Exercises  
**1. Run "ggplot(data = mpg)" what do you see?**
```{r}
ggplot(data = mpg)
```

This renders a blank coordinate plane. The graphing layer and mapping were not defined.  

**2. How many rows are in mtcars? How many columns?**

There are `r nrow(mtcars)` rows and `r ncol(mtcars)` columns in mtcars. (inline coding used)  

**3. What does the drv variable describe? Read the help for ?mpg to find out.**
```{r}
?mpg
```

According to the mpg help file drv describes the drive of the car such that "f = front-wheel drive, r = rear wheel drive, 4 = 4wd".  

**4. Make a scatterplot of hwy vs cyl.**
```{r}
ggplot (data = mpg) +
  geom_point(mapping = aes(x = cyl, y = hwy))
```

**5. What happens if you make a scatterplot of class vs drv. Why is the plot not useful?**
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = drv, y = class))
```
Drive and class are two fields that are used to describe the data, therefore they do not influence one another and will not yield useful insight when compared. 




### 3.3 Aesthetic mappings

#### 3.3 Exercises
**1. What’s gone wrong with this code? Why are the points not blue?**
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
```

In order to make change the color of the points to blue, the aesthetic value color must appear _outside_ of the aesthetic parens, as show below.

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
```

**2. Which variables in mpg are categorical? Which variables are continuous? (Hint: type ?mpg to read the documentation for the dataset). How can you see this information when you run mpg?**  
To determine which variables in mpg are continuous and which are categorical we can examine the summary of mpg
```{r}
summary(mpg)
```
The variables manufacturer, model, trans, drv, fl, and class are categorical.
The variables displ, year, cly, cty, and hwy are continuous. 

**3. Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables?**
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = cty))
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, size = cty))
#ggplot(data = mpg) + 
  #geom_point(mapping = aes(x = displ, y = hwy, shape = cty))
# will not plot

```
Continuous data will be scaled on a gradient for color and size but not plot if assigned to shape. 

**4. What happens if you map the same variable to multiple aesthetics?**
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = cyl, size = cyl))
```
The variable will be plotted using both aesthetics.

**5. What does the stroke aesthetic do? What shapes does it work with?**
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, stroke = 1))
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, stroke = 5))
```
The stroke aesthetic changes the size of the shapes in geom_point. It seems that stroke will apply to all values for shape between 1-25. 

**6. What happens if you map an aesthetic to something other than a variable name, like aes(colour = displ < 5)?**
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = displ < 5))
```
Setting an aesthetic to a value besides a variable name will plot the aesthetic accoridng to that qualifier. In the example above color is plotted between two classes, data where displ >5 and < 5. 


## 3.5 Facets
### 3.5 Exercises



