---
title: "mpg.practice"
author: "Veronica"
date: "4/24/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## MPG dataset practice
#### 3.2 First Steps

```{r}
mpg
```


```{r}
ggplot(data = mpg) +
  geom_point(mapping =aes(x = displ, y = hwy))
```
##### 3.2 Exercises
**1. Run "ggplot(data = mpg)" what do you see?**
```{r}
ggplot(data = mpg)
```

This renders a blank coordinate plane. The mapping and graphing layer were not defined.  

**2. How many rows are in mtcars? How many columns?**

There are `r nrow(mtcars)` rows and `r ncol(mtcars)` columns in mtcars.  

**3. What does the drv variable describe? Read the help for ?mpg to find out.**
```{r}
?mpg
```

according to the mpg help file drv described the drive of the car such that "f = front-wheel drive, r = rear wheel drive, 4 = 4wd".  

**4. Make a scatterplot of hwy vs cyl.**
```{r}
ggplot (data = mpg) +
  geom_point(mapping = aes(x = cyl, y = hwy))
```

**5. What happens if you make a scatterplot of class vs drv. Why is the plot not useful?**
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = drv, y = class))
```
Drive and class are two fields that are used to describe the data, therefore they do not influence one another and will not yeild useful data when compared. 



#### 3.3 Aesthetic mappings

