---
title: "Relational_data"
author: "Veronica"
date: "6/23/2017"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nycflights13)
library(Lahman)
library(babynames)
library(nasaweather)
library(fueleconomy)
library(maps)
```

### 13.3 Keys

**1. Add a surrogate key to flights.**  
```{r}
flights %>% 
  mutate(ID = row_number()) %>% 
  summary()
```

**2. Identify the keys in the following datasets:**  
```{r}
Lahman::Batting %>% 
  count(playerID, yearID, stint) %>% 
  filter(n>1)

babynames2 <- as.data.frame(babynames)

babynames2 %>% 
  count(year, sex, name)  %>% 
  summary()
#this data set will not filter

nasaweather::atmos %>% 
  count(lat, long, year, month) %>% 
  filter(n>1)

fueleconomy::vehicles %>% 
  count(id) %>% 
  filter(n>1)

#ggplot2::diamonds 
#There is no key
```

**3. Draw a diagram illustrating the connections between the Batting, Master, and Salaries tables in the Lahman package. Draw another diagram that shows the relationship between Master, Managers, AwardsManagers.**

**How would you characterise the relationship between the Batting, Pitching, and Fielding tables?**

Batting decribed offensive performance of a player while Pitching and Fielding describe the defensive performance.

### 13.4 Mutating Joins  

```{r}
flights2 <- flights %>% 
  select(year:day, hour, origin, dest, tailnum, carrier)
```

**1. Compute the average delay by destination, then join on the airports data frame so you can show the spatial distribution of delays. Here’s an easy way to draw a map of the United States:**

```{r}
flights %>% 
  select(dest, arr_delay) %>% 
  group_by(dest) %>% 
  summarize(ave.delay = mean(arr_delay, na.rm =T)) %>%
  right_join(airports, c("dest" = "faa")) %>%
  ggplot(aes(lon, lat, color = ave.delay)) +
    borders("state") +
    geom_point() +
    coord_quickmap()
```

not sure what my outliers here are

**2. Add the location of the origin and destination (i.e. the lat and lon) to flights.**
```{r}
airports2 <- airports %>% 
  select(faa, lat, lon)

flights2 %>% 
  left_join(airports2, c("origin" = "faa")) %>% 
  rename("origin_lat" = lat, "origin_lon" = lon) %>% 
  left_join(airports2, c("dest" = "faa")) %>% 
  rename("dest_lat" = lat, "dest_lon" = lon)
```

**3. Is there a relationship between the age of a plane and its delays?**
```{r}
planes2 <- planes %>% 
  select(tailnum, year)

flights %>% 
  group_by(tailnum) %>% 
  summarize(ave.delay = mean(arr_delay, na.rm =T)) %>% 
  left_join(planes) %>% 
  ggplot(aes(year, ave.delay)) +
  geom_point()
```

There does not appear to be a coorelation between the year the plane was made and ave.delay for that tail number.

**4. What weather conditions make it more likely to see a delay?**
```{r}
flights %>% 
  left_join(weather) %>% 
  filter(wind_speed <= 100) %>% 
  ggplot(aes(wind_speed, dep_delay)) +
  geom_point()

flights %>% 
  left_join(weather) %>% 
  ggplot(aes(precip, dep_delay)) +
  geom_point()

flights %>% 
  left_join(weather) %>% 
  ggplot(aes(visib, dep_delay)) +
  geom_point()

flights %>% 
  left_join(weather) %>% 
  ggplot(aes(temp, dep_delay)) +
  geom_point()

flights %>% 
  left_join(weather) %>% 
  filter(wind_gust <= 100) %>% 
  ggplot(aes(wind_gust, dep_delay)) +
  geom_point()
```

**5. What happened on June 13 2013? Display the spatial pattern of delays, and then use Google to cross-reference with the weather.**

```{r}
flights %>%
  filter(month == 6, day == 13, year == 2013) %>%
  ggplot(aes(dep_time, dep_delay)) +
  geom_point()
```

There was an unusually strong summer storm that passed through New York in the afternoon.

### Filtering Joins

**1. What does it mean for a flight to have a missing tailnum? What do the tail numbers that don’t have a matching record in planes have in common? (Hint: one variable explains ~90% of the problems.)**

```{r}
flights %>% 
  filter(tailnum == NA)
```





