---
title: "Relational_data"
author: "Veronica"
date: "6/23/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nycflights13)
```

### 13.3 Keys

**1. Add a surrogate key to flights.**  
```{r}
flights %>% 
  mutate(ID = row_number()) %>% 
  summary()
```

**2. Identify the keys in the following datasets:**  
```{r}
Lahman::Batting %>% 
  count(playerID, yearID, stint) %>% 
  filter(n>1)

#babynames::babynames %>% 
#  count(year, sex, name) %>% 
#  filter(n>1)

nasaweather::atmos %>% 
  count(lat, long, year, month) %>% 
  filter(n>1)

fueleconomy::vehicles %>% 
  count(id) %>% 
  filter(n>1)

#ggplot2::diamonds 
#There is no key
```

**3. Draw a diagram illustrating the connections between the Batting, Master, and Salaries tables in the Lahman package. Draw another diagram that shows the relationship between Master, Managers, AwardsManagers.**

**How would you characterise the relationship between the Batting, Pitching, and Fielding tables?**


 



